--- 
title: "Stat 210 Project" 
author: "Sofia Silvosa" 
format: pdf 
---

# Introduction 



```{r read-data, message = F, warning = F, echo=FALSE}
library(tidyverse)
library(ggfortify)
library(viridis)
library(tidymodels)
library(ggplot2)
library(broom.mixed)
library(car)
library(leaps)
library(broom)
library(knitr)
library(kableExtra)
library(dplyr)
library(MASS)
library(lme4)
library(glmnet)
library(readr)
library(readr)
movies <- read_csv("movies.csv")

```

*Background and Significance:*
In the entertainment sphere,there has been recent discussion on adequate
female representation in film. Many in the film industry have recently praised
the increased representation of women in film (Nolfi,2017).For a while Hollywood
produced very little movies with female protagonists; let alone movies with
appropriate representation of women (Goodman, 2017).
However, merely having female
protagonists does not paint the full picture if
whether a movie actually does a good job in female representation. Many have 
used the so-called Bechdel test to make this judgement. This test
was developed by Allison Bechdel in 1985 and has recently become a digital
sensation (Hickey, 2014).The Bechdel test is a simple test which deems that
a movie has adequate 
female character representation in their film if it passes the following three
criteria:(1) it has to have at least two women in it, who (2) who talk 
to each other, 
about (3) something besides a man (https://bechdeltest.com/). Many have argued 
that there has been significant increase of female 
representation throughout the decades thanks to increased diversity 
initiatives and more women at the helm of the film industry 
(UCLA-Hollywood-Diversity-Report-2022-Film). 

*Research Question & Hypothesis*:
We're interested in evaluating what variables are important in
predicting whether a movie passes or fails
the Bechdel test. We are interested, specifically, if the time period
in which a movie was released predicts whether a movie passes the Bechdel test.
In other words, does the decade a movie was released predict whether it passes
the Bechdel test while controlling for other pivotal variables in our model? 
As shown by recent news and media, more modern movies seem to have more
increased representation of women in their films. Thus, we predict that 
the time period in which a movie was released will play a significant role
in whether the movie fails or passes the Bechdel test. More specifically,
we predict that movies released in the later decades of our model 
will have a higher probability of passing the Bechdel test than
those from earlier decades.

*Data*:
To explore our research question, we'll be using a data set used in the
FiveThirtyEight story titled “The Dollar-And-Cents Case Against Hollywood’s
Exclusion of Women” (Hickey, 2014). The data set includes observations from 
1794 films that were released between 1970-2013.The data set was organized by 
combiningdata from two major sources. One of them came from the BechdelTest.com:
a website operated by committed moviegoers who analyze films and 
ascertain if they pass the Bechdel test.To provide financial 
information for the chosen films, the FiveThirtyEight team gathered data 
from the website The-Numbers.com, a leading site for box office and budget 
data. The finalized *movies.cv* data set includes information
detailing the title of the film, the year it was released, its domestic gross, 
budget and international gross (both accounting for inflation at the time 
of data collection and without). See Data Dictionary for more details.
Furthermore, for our purposes, the data set also includes two important 
columns regarding whether the movie's performance on the Bechdel test. 
The column "binary" specifically states whether the movie passed or 
failed the Bechdel test in a binary fashion. The column "clean_test" goes a
bit more in detail, regarding
how the film failed the Bechdel test or if it was unclear whether the movie 
passed or not. The clean test variable has five levels: ok (Passed), no women
(No women in the film), dubious (unclear result), no talk(women did not
talk to each other), and men (women only talked about men). We created new 
variables for the purposes of our analysis. For starters, we created a decades
variable that detailed which decade the given film was released in. 
The variable ended up having 5 different levels: 1970s, 1980s, 1990s, 2000s and
2010s. We also created a new variable titled "passfail"
which is essentially the same as our 'binary' variable but instead uses dummy
values to illustrated whether the movies passed (passfail=1) failed 
(passfail=0) the Bechdel test. We excluded 18 observations from our final
analysis. These observations had no values for their total domestic gross 
and total budget. These are variables of interest that
we want to control for. Since we still had 1776 observations left in our
data set, our statistical analyses will not be greatly affected by the removal
of these observations.

*Exploratory Data Analysis*
```{r adding-new-variables, message = F, warning = F, echo= FALSE}

# rename columns


colnames(movies)[which(names(movies) == "domgross_2013$")] <- "dom_Gross2013"
colnames(movies)[which(names(movies) == "budget_2013$")] <- "budget_Gross2013"
colnames(movies)[which(names(movies) == "intgross_2013$")] <- "int_Gross2013"

#remove NAs


movies1 <- movies[!(movies$dom_Gross2013=="#N/A" | movies$budget_Gross2013 =="#N/A"),]

movies1$ dom_Gross2013<- as.numeric(movies1$dom_Gross2013)
movies1$ budget_Gross2013 <- as.numeric(movies1$budget_Gross2013)
movies1$ int_Gross2013<- as.numeric(movies1$int_Gross2013)

#change units to a million

movies2 <- movies1 %>% 
  mutate(dom_gross_mil = dom_Gross2013 /1000000) %>% 
  mutate(int_gross_mil = int_Gross2013 /1000000) %>% 
    mutate(  budget_mil = budget_Gross2013 /1000000)


movies_decades <- movies2 %>% 
  mutate(decade= case_when(year < 1980 ~ '1970s' ,
                         1980<= year & 1990 > year ~ '1980s' ,
                         1990<= year & 2000 > year ~ '1990s' ,
                         2000<= year & 2010 > year ~ '2000s ',
                       2010 <= year ~ '2010s ' ))  
  
  ggplot(data= movies_decades, mapping= aes(x=decade, fill=binary)) +
  geom_bar(position = "fill") +
labs(title="The % of Films That pass the Bechdel Test has Increased", 
       subtitle = "The 2000s had the largest percentage of movies that passed
     the Bechdel Test",
     y= 'Propportion of Bechdel Test Results', fill= 'Bechdel Test Results',
     x= 'Decade')


table <- movies_decades %>% 
  group_by(decade) %>% 
  summarise(number_of_movies =n()) 


table2 <- movies_decades %>%
  group_by(decade) %>%
  summarize(passed_movies = sum(binary == 'PASS'))



proportion_70s = 14/ 54
proportion_80s = 36 / 125
proportion_90s = 147/ 337
proportion_2000s = 409/ 840
proportion_2010s = 197/ 438




 



```
We can see above that there is general upward trend across
the decades with an increased percentage of movies that pass the 
Bechdel test. We see, specifically that the movies from our data set
that premiered in the 2000s, about 48% of the films passed the Bechdel test.
The 2010s performed in a similar fashion, with 45% of the films released
passing the Bechdel test (it is important to note however that this
data set only includes movies till 2013, thus it does not paint the
full picture of female representation in film from this decade) The decade 
with the lowest percentage of movies that passed the Bechdel test was the 
1970s, withonly 25% passing. Below we also see a similar trend when looking
years as our independent variable and with clean test as our dependent variable,
revealing that films released in the 21st century have a higher percentage
of movies that pass the Bechdel test. Moreover, we also see that "no talk"
talk seems to be the most common reason for a movie to fail the Bechdel test.

```{r graph, message = F, warning = F, echo= FALSE}


movies1_0 <- movies_decades %>% 
 mutate(passfail= ifelse(binary =='PASS', 1, 0))
  
  ggplot(data= movies_decades, mapping= aes(x= year, fill=clean_test)) +
  geom_bar()+
  labs(title="The % of films that pass the Bechdel Test increases through time",
  subtitle = "No Talk seems to be the most common reason movies fail the Test",
     y= 'Number of Movies', fill= 'Bechdel Test Results',
     x= 'Year')




```
# Methodolody
We are interested in running a logistic regression model in order to evaluate
whether the time period a given film was released in is associated with its
scoreon the Bechdel test. We believe this is an appropriate model,
firstly, because our data passes the independence assumption. We can
assume that independence is
met because our observations are most likely not correlated with each other. 
Each of our movie titles are independent from each other and knowing something 
about one of our observations does not reveal anything substantial about another 
observation. For our outcome variable, we chose to use the binary 
version detailing whether given film's performance on the Bechdel test rather
than our "cleantest" variable because we felt that a binary outcome variable 
would generate a simpler model. Furthermore, if we ran a logistic regression
model with the "cleantest" variable as our outcome variable, 
we would have to use a multinomial regression model to test out our
research question. A multinomial regression model would not make sense 
in this context since the independence of irrelevant alternatives assumption 
would be violated. This assumption assumes
that, in a multinomial logistic regression model, the relative odds of choosing 
one option over another should not be influenced
by the inclusion or exclusion of an additional option. This does not make
sense since the inclusion or exclusion 
of a Bechdel test result category could have an effect on our final analysis. 
For example, if a given film with plentiful female representation that was 
released in 2013 (which according to our hypothesis means it has a greater
chance of passing the Bechdel test) was included our model but the only
two categories taken accounted for whether "notalk" and "dubious," 
our model would predict it was it fit the 'dubious' category. However, 
if the "ok" was included in the mix, this would change our
predictive probability.

Next, we evaluated whether to use the "year" or "decade" variable for
our investigation. To decide which variable to use, we 
checked whether our continuous variable
"year" passed the linearity assumption.


```{r log-linearity, message = F, warning = F, echo= FALSE, out.width="70%"}

 
  library(Stat2Data)
emplogitplot1(passfail ~ year,
              data = movies1_0,
              ngroups = 10)



```
As shown above, the points are not evenly scattered through our line of best
fit, therefore we have decided that this is not an appropriate variable to use
in our model and instead will use the decade variable.
We could have transformed the "year" variable, quadratically for example
in order to pass this linearity assumption. However, this will complicate
our interpretation of our model gravely since our outcome variable of
interest would be transformed. Thus, for the purposes of our 
investigation, we will use the 'decades' categorical variable. 
For our other predictor variables,we are interested in using international
gross, domestic gross, budget and an interaction between international gross 
and domestic gross. These variables have been shown to affect Bechdel test
results in the past (Hickey, 2014) thus we thought it would be smart to control
for these variables in our model.Furthermore, we chose to introduce an
interaction term between domestic 
gross and international gross because these two variables tend to depend on
each other--when a movie does well internationally this depends on its domestic
success since movie industries usually stop investing in marketing for
the film if the film performs poorly domestically first since the film 
could be seen as a "lost cause" (stephenfollows.com).

We also checked to see if the linearity condition was met for our other
continuous predictors (see Appendix for linearity plots for all
continuous predictors). 
Only the budget variable passed the linearity
assumption. The linearity condition was not met for our predictors
dom_Gross2013 and
Int_Gross2013. In order to deal with this violation of our linearity assumption,
we applied
a log transformation to these variables.

This is our proposed model that we will use to investigate our research question:

Our model is as follows. For each decade the film premiered in $i$,


$p / (1- p)$ = Odds of passing the Bechdel Test

$log(p / (1-p) =\beta_0 + \beta_1(Budget)_i + \beta_2(log(IntGross))_i + \beta_3(log(DomGross) + \beta_4I(decade1980s)_i  +   \beta_5I(decade1990s)_i  +   \beta_6I(decade2000s)_i     + \beta_7I(decade2010s)_i   +  \beta_8(log(IntGross))_i * log(DomGross)_i$


## Results

*Hypothesis Test*
We will utilize logistic regression model using the final model discussed 
in the previous section
to see if the decade in which a movie was released
predicts whether a movie passes or fails the Bechdel test, while controlling 
for other variables in our model. We will first run a hypothesis test to 
see if there sufficient statistical evidence to suggest that there is
a relationship between the decade in which a movie is released and
whether the given film passes the Bechdel test. 

 
We will run a hypothesis at the $a$ = 0.05 level.


Null hypothesis: 
There is not sufficient evidence to suggest that the decade in 
which a movie
premiered is associated with the differential odds of the 
passing the Bechdel test, while
controlling for all of the variables represented in our model.

$H_0$ : All of our $\beta$ terms for decade ( $\beta_4$, $\beta_5$, $\beta_6$ 
and $\beta_7$ ) are equal to zero.


Alternative Hypothesis: There is sufficient evidence to suggest
that the decade in which a movie premiered is associated with the differential 
odds of passing the Bechdel test, while
controlling for all of the variables represented in our model.

$H_1$: At least one of our $\beta$ terms for
decade ( $\beta_4$, $\beta_5$, $\beta_6$ 
or $\beta_7$ ) is not equal to zero.



```{r anova, message = F, warning = F, echo= FALSE, width= '70%'}


  
m22 <- glm(passfail ~ decade+ budget_Gross2013 +log(dom_Gross2013) +
             log(int_Gross2013) + log(int_Gross2013) * log(dom_Gross2013),
          data = movies1_0,
          family = "binomial")


  
m21 <- glm(passfail ~  budget_Gross2013 + log(dom_Gross2013) +
             log(int_Gross2013) + log(int_Gross2013) * log(dom_Gross2013),
          data = movies1_0,
          family = "binomial")

anova(m22, m21, test= 'Chisq') %>% 
  tidy() 

qf(p= 0.0002963195, 4 , 1776, lower.tail= FALSE)


```
We will be conducing an F test for this formal hypothesis test. 
The F statistic is 5.31495 and falls under an F distribution with 4
numerators of degrees of freedom and 1776 denominator degrees of freedom.
We reject the null hypothesis in this case since our p value is less than 
0.05, meaning that at least one of our decade predictor levels has a slope
that is not 0. There is sufficient evidence to suggest that the decade in
which a movie premiered in is associated with the differential odds of 
passing the Bechdel test, while controlling for all of the variables listed in
the previous section (and adjusting for log(domGross), log(intgross) 
and log(intgross)* log(domGross.)


```{r glm-model, message = F, warning = F, echo=FALSE}



m22 <- glm(passfail ~ decade+ budget_Gross2013 +log(dom_Gross2013) +
             log(int_Gross2013) + log(int_Gross2013) * log(dom_Gross2013),
          data = movies1_0,
          family = "binomial")  


summary(m22)$coef %>% 
  kable()





```



*Coefficient Interpretations*

While controlling for the variables in our model, the odds of passing the
Bechdel test for a movie that premiered in the 1990s is estimated to be 2.3379
times the odds of a film 
that was released in the 1970s (adjusting for our transformed variables). 
While controlling for the variables in our model, the odds of passing the
Bechdel test for a movie that premiered in the 2000s is estimated to be 2.959
times the odds of a film 
that was released in the 1970s (adjusting for our transformed variables).
While controlling for the variables in our model, the odds of passing the
Bechdel test for a movie that premiered in the 2010s is estimated to be 2.601
times the odds of a film 
that was released in the 1970s (adjusting for our transformed variables).

*Predictive Power*:
Even though our result was significant, we also want to check for 
the predictive power of our model. We will thus check the obtain predicted
probabilities of success for our binary variable "passfail." We will impose
a threshold of 0.5 as our classifier. We will check the sensibility and
sensitivity for our model to asses its predictive power and construct an ROC
curve to show how specificity and specificity change as our
discrimination threshold changes (see ROC curve in Appendix).
Calculating the area underneath the ROC curve will also give us more insight
into the predictive power of our model.


```{r prediction-power, message = F, warning = F, echo=FALSE}


m2_aug <- augment(m22)
m2_aug <- m2_aug %>%
mutate(prob = exp(.fitted)/(1 + exp(.fitted)), pred_leg = ifelse(prob >
0.5,
'Pass',
'Fail')) 

m2_aug %>% 
  roc_auc(
    truth = as.factor(passfail),
    prob, 
    event_level = "second"
  ) %>% 
  kable()
  


 


```
Our Area under the curve is 0.62 which is higher than 0.05--meaning 
that our model suggests some discrimination its outcomes are not completely
governed by chance.


```{r prediction-power-ss, message = F, warning = F, echo=FALSE}


  
table(m2_aug$pred_leg, m2_aug$passfail) %>% 
  kable()

 


```

As shown by our table above, our specificity was 67.82%. Furthermore, our 
sensitivity was 50.13%. 

## Discussion ##

Ultimately, our model demonstrates that there is an association between
the decade in which a film was released and whether it passes or fails the
Bechdel test. This was illustrated through our hypothesis test
which demonstrated a significant association between the decade in which
a film was released and its score on the Bechdel test. Furthermore,
our logistic regression model demonstrated that if a movie premiered
during the 2010s, 2000s or the 1990s, it had higher odds of passing
the Bechdel test compared to a movie that came out in the 1970s.

We can conclude, therefore, that the time period in which a movie 
came out in does in fact hold predictive power in predicting whether
the given film passed or failed the Bechdel test. Perhaps the movie
industry has indeed progressed throughout time and has released films 
with more adequate female representation in film. This could be a 
result of the recent diversity campaigns in Hollywood and a result 
of an increase of women working in the film industry, allowing for 
more adequate representation for female characters. 
However, the predictive power of our model was, overall, pretty weak. 
Our model only detected 50% of the films that in actuality passed the 
Bechdel test as shown by our sensitivity calculation. Our model did a 
slightly better job at predicting which movies failed the Bechdel
test as shown by the fact that our model predicted 68% of films that
in actuality failed the Bechdel test. 

Our model most likely did not obtain enough predictive power
because of the limitations our investigation faced. Firstly,
our model had an uneven amount of films that premiered in our decades
of interest. We had 54 observations for the 1970s, 125 observations for
1980s, 337 for 1990s, 840 for the 2000s and 438 for the 2010s (see Appendix 
for summary statistics). This could have, in turn, affected our statistical 
inference 
since there was an unbalance of observations for each level of our predictor 
variables. Another limitation is the fact some of the coefficients
in our model are pretty difficult to interpret due to our inclusion of 
transformed variables.Lastly, another limitation includes the lack of variables
in our data set. Our data set only included 15 variables to begin with
which have us a limited amount of variables to choose from and to control 
for in our final model. There are many future directions that could be taken
with this investigation. For starters, it could be interesting to have a dataset
that represents moredecades (from the 1920s till 2020s). Furthermore,
future investigationsshould include more variables in their model. 
It could be interestingto see if the number of women working on the film
is associated withthe film's score on the Bechdel test. 
Film genre could also be an interesting variable to look at; it is possible
that movies from certain genres are predicted to have higher odds of passing 
the Bechdel test than others. 

## Sources

https://socialsciences.ucla.edu/wp-content/uploads/2022/03/UCLA-Hollywood-
Diversity-Report-2022-Film-3-24-2022.pdf

https://www.nytimes.com/2017/02/21/movies/women-protagonists-movies-2016.html
https://www.tandfonline.com/doi/pdf/10.1080/14680777.2016.1234239?
needAccess=true

https://bechdeltest.com/

https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-
hollywoods-exclusion-of-women/

https://ew.com/oscars/2017/07/03/oscars-2017-new-members-diversity-push/

https://github.com/fivethirtyeight/data/tree/master/bechdel

https://stephenfollows.com/important-international-box-office-hollywood/


## Appendix ##

*Additional Exploratory Analysis*


```{r how-many-movies-passed-test-across-decades22, message = F, warning = F, echo= FALSE}


  
  ggplot(data= movies_decades, mapping= aes(x= binary, y= dom_Gross2013)) +
  geom_boxplot()  +
  labs(title= "Association between Bechdel Test Result and Domestic Gross",
subtitle= 'Movies that Fail the Bechdel Test perform better in the Box Office',
x= 'Bechdel Test Result', y= 'Domestic Gross')
  

  ggplot(data= movies_decades, mapping= aes(x=decade, fill=binary)) +
  geom_bar() +
labs(title="More Movies Have Passed the Bechdel Test Over Time", 
       subtitle = "The current data set overrepresents 2000s movies",
     y= 'Number of Films', fill= 'Bechdel Test Results',
     x= 'Decade')




```

*Linearity Plots for Continous Predictors*

```{r linearity-plots-appendix, message = F, warning = F, echo= FALSE}


  library(Stat2Data)
emplogitplot1(passfail ~ (int_Gross2013),
              data = movies1_0,
              ngroups = 10)

 

  library(Stat2Data)
emplogitplot1(passfail ~ dom_Gross2013,
              data = movies1_0,
              ngroups = 10)


```

```{r lin-assumption, message = F, warning = F, echo=FALSE}


 
library(Stat2Data)
emplogitplot1(passfail ~ log(dom_Gross2013),
              data = movies1_0,
              ngroups = 10)



  library(Stat2Data)
emplogitplot1(passfail ~ budget_Gross2013,
              data = movies1_0,
              ngroups = 10)




 library(Stat2Data)
emplogitplot1(passfail ~ log(int_Gross2013),
              data = movies1_0,
              ngroups = 10)


 library(Stat2Data)
emplogitplot1(passfail ~ (log(dom_Gross2013) * log(int_Gross2013)),
              data = movies1_0,
              ngroups = 10)
 






# nice
```


*ROC Curve*



```{r prediction-power-ROC, message = F, warning = F, echo=FALSE}


m2_aug %>% 
  roc_curve(
    truth = as.factor(passfail),
    prob, 
    event_level = "second"
  ) %>% 
  autoplot() 

```
